Turkey Hemorrhagic Enteritis Virus transcriptome profiling

### 2023-05-30 manuscript_thev_transcriptome.Rmd
Modify stringtie to create count tables modify putative genomic map to show region labels update some text in manuscript

 Virulent strains (THEV-V) and avirulent strains (THEV-A) of THEV are serologically indistinguishable, infecting turkeys, chickens, and pheasants. They infect via the fecal-oral route, and the THEV-V cause different clinical diseases in these birds [@Harrach2008; @Beach2006]. In turkeys, the THEV-V cause hemorrhagic enteritis (HE), a debilitating acute disease affecting predominantly 6-12-week-old turkeys characterized by immunosuppression (IS), weight loss, intestinal lesions leading to bloody diarrhea, splenomegaly, and up to 60% mortality [@Beach2006; @Gross1967]. HE is the most economically significant disease caused by any strain of THEV [@Beach2006]. While the current vaccine strain (a THEV-A isolated from a pheasant, Virginia Avirulent Strain [VAS]) have proven effective at preventing HE in young turkey poults, it still retains the immunosuppressive ability. Thus, vaccinated birds are rendered more susceptible to opportunistic infections and death than unvaccinated cohorts leading to substantial economic losses [@Beach2006; @Rautenschlein2000; @Larsen1985; @Dhama2017]. The induced IS also interferes with vaccination schemes for other infections of turkeys [@Beach2006; @Rautenschlein2000]. To eliminate this immunossupressive side-effect of the vaccine, a thorough investigation of the culprit viral factors (genes) mediating this phenomenon is essential. However, the transcriptome (splicing and gene expression patterns) of THEV has not been characterized, making the investigation of specific viral genes for possible roles in causing IS impractical. To identify such viral factors, a well-characterized transcriptome of THEV is required to set the stage for experimentation with specific viral genes that may mediate IS.

@@

## TABLES AND FIGURES     
![Fig.1](results/r/thev_genomic_map.png)  
**Figure 1. _Genomic map of THEV avirulent strain_**. The central horizontal line represents the double-stranded DNA marked at 5kb intervals as white line breaks. Blocks represent viral genes. Blocks above the DNA line are transcribed rightward, those below are transcribed leftward. pTP, DBP and 33K predicted to be spliced are shown as having tails. Shaded regions indicates regions containing genes (colored red) that are not conserved across adenovirus genera (i.e. "genus-specific"). Genes colored in blue are "genus-common".  Gene colored in light green is conserved in all but Atadenoviruses. The U exon (light blue) is an incomplete gene present in almost all AdVs.

>>>>

 Virulent strains (THEV-V) and avirulent strains (THEV-A) of THEV are serologically indistinguishable, infecting turkeys, chickens, and pheasants. They infect via the fecal-oral route, and the THEV-V cause different clinical diseases in these birds [@Harrach2008; @Beach2006]. In turkeys, the THEV-V cause hemorrhagic enteritis (HE), a debilitating acute disease affecting predominantly 6-12-week-old turkeys characterized by immunosuppression (IS), weight loss, intestinal lesions leading to bloody diarrhea, splenomegaly, and up to 80% mortality [@Beach2006; @Beach2009a; @Gross1967]. HE is the most economically significant disease caused by any strain of THEV [@Beach2006]. While the current vaccine strain (a THEV-A isolated from a pheasant, Virginia Avirulent Strain [VAS]) have proven effective at preventing HE in young turkey poults, it still retains the immunosuppressive ability. Thus, vaccinated birds are rendered more susceptible to opportunistic infections and death than unvaccinated cohorts leading to substantial economic losses [@Beach2006; @Rautenschlein2000; @Larsen1985; @Dhama2017]. The induced IS also interferes with vaccination schemes for other infections of turkeys [@Beach2006; @Rautenschlein2000]. To eliminate this immunossupressive side-effect of the vaccine, a thorough investigation of the culprit viral factors (genes) mediating this phenomenon is essential. However, the transcriptome (splicing and gene expression patterns) of THEV has not been characterized, making the investigation of specific viral genes for possible roles in causing IS impractical. To identify such viral factors, a well-characterized transcriptome of THEV is required to set the stage for experimentation with specific viral genes that may mediate IS.

@@

## TABLES AND FIGURES     
![Fig.1](results/thev_genomic_map.png)  
**Figure 1. _Genomic map of THEV avirulent strain_**. The central horizontal line represents the double-stranded DNA marked at 5kb intervals as white line breaks. Blocks represent viral genes. Blocks above the DNA line are transcribed rightward, those below are transcribed leftward. pTP, DBP and 33K predicted to be spliced are shown as having tails. Shaded regions indicates regions containing "genus-specific" genes (colored red). Genes colored in blue are "genus-common".  Gene colored in light green is conserved in all but Atadenoviruses. The UXP (light blue) is an incomplete gene present in almost all AdVs. XXXXX add text to expalain region labels here or in the intro.
